The Legacy of the European Post-Master in Urbanism at TU Delft: A Text Mining Approach

### 2023-12-06 analysis/paper/paper.qmd
Update column names in topic summary table
Update topic model description and fix k to 5

Similar to clustering on numeric data, topic modelling is a method of unsupervised classification of topics found in a collection of documents [@silge2017]. We used the Latent Dirichlet Allocation (LDA) algorithm for topic modeling. In LDA, each document is a mixture of topics and each topic is a mixture of words. The probability of a term being part of a topic is given by the $\beta$ (beta) statistic, while the probability of a document being part of a topic is given by the $\theta$ (theta) statistic.

LDA is a Bayesian generative topic model which assumes that each topic is a distribution over words and each document is a distribution over topics [@blei2012; @grün2011; @blei2003].

Latent topics are measured as patterns of word co-occurrence

Describe model parameters:

-   collection of M = `r nrow(emu)` documents

-   corpus D

-   N = `r length(emu_words$word)` words

-   vocabulary of V = `r length(unique(emu_words$word))` words

Determining tuning parameters:

-   The number of topics was qualitatively determined
-   Evaluating the number of topics quantitatively can be done through the measure of **topic coherence**
    -   Calculate topic coherence for K values ranging from 2 to K
    -   Plot K values against coherence scores and select K with the highest coherence value

A K with a smaller value will yield more general topics whereas a K with a high value will result in more specific topics. Hence, finding a suitable value entails a trade-off between interpretability and specificity. This is very much dependent on the data at hand and the research question.

#### Determining the value of K empirically

```{r topic-model-lda-1}
# Determine K value empirically
k = get_number_of_topics(emu_dtm)
```

Fit model with K determined in previous step...

```{r topic-model-lda-1}
emu_lda_k <- LDA(emu_dtm, k = k, method = "Gibbs",
               control = list(seed = 2023, iter = 500, verbose = 25))

beta_k <- posterior(emu_lda_k)$terms
theta_k <- posterior(emu_lda_k)$topics

# Add pseudo-names to topics based on the top 5 words in each topic
topic_names_k <- name_topics(beta_k)
```

@@

Run a cluster dendrogram (Hellinger distance) to see which topics should be grouped together

```{r}
# Perform hierarchical cluster analysis.

rownames(beta_k) <- topic_names_k
dend <- as.dist(beta_k) |> 
  hclust() |> 
  as.dendrogram()

# Plot with Color in function of the cluster
leafcolor <- colormap::colormap(colormap = colormap::colormaps$viridis, nshades = 5, format = "hex", alpha = 1, reverse = FALSE)
dend |> 
  # set("labels_col", value = leafcolor, k=5) %>%
  # set("branches_k_color", value = leafcolor, k = 5) %>%
  # plot(horiz=TRUE, axes=FALSE)
  ggdendro::ggdendrogram(rotate = TRUE)

dendr    <- ggdendro::dendro_data(dend, type="rectangle") # convert for ggplot
clust    <- cutree(as.dist(beta_k) |> hclust(),k = 5)                    # find 2 clusters
clust.df <- data.frame(label=names(clust), cluster=factor(clust))
# dendr[["labels"]] has the labels, merge with clust.df based on label column
dendr[["labels"]] <- merge(dendr[["labels"]],clust.df, by="label")
# plot the dendrogram; note use of color=cluster in geom_text(...)
ggplot() + 
  geom_segment(data=ggdendro::segment(dendr), aes(x=x, y=y, xend=xend, yend=yend)) + 
  geom_text(data=ggdendro::label(dendr), aes(x, y - 0.002, label=label, hjust=0, color=cluster), 
           size=3) +
  coord_flip() + 
  scale_y_reverse(expand=c(0.1, 0.1)) +
  theme(axis.line.y=element_blank(),
        axis.ticks.x=element_blank(),
        axis.ticks.y=element_blank(),
        axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        panel.background=element_rect(fill="white"),
        panel.grid=element_blank())
```

@@

#### Determining the value of K qualitatively

OR Use LDA clustering with 3 classes (corresponding to the three exchange universities) to see how well the clustering is aligned with where the student carried out their exchange semester. Use a confusion matrix to visualize the relation between real values and values assigned through clustering.

```{r topic-model-lda-1}
## Determine K value qualitatively
## K = 5 corresponds to the 2 TUD semesters and the 3 exchange semesters combined 
emu_lda_5 <- LDA(emu_dtm, k = 5, method="Gibbs",
               control = list(seed = 2023, iter = 500, verbose = 25))

@@

topic_proportions |> 
  left_join(counts_of_primary_topics, by = "topic")
```

>>>>

Similar to clustering on numeric data, topic modelling is a method of unsupervised classification of topics found in a collection of documents [@silge2017]. We used the Latent Dirichlet Allocation (LDA) algorithm for topic modeling. LDA is a Bayesian generative topic model which assumes that each topic is a distribution over words and each document is a distribution over topics [@blei2012; @grün2011; @blei2003]. This means that each document is a mixture of topics and each topic is a mixture of words. The probability of a term being part of a topic is given by the $\beta$ (beta) statistic, while the probability of a document being part of a topic is given by the $\theta$ (theta) statistic.

The corpus (D) consists of a collection of M = `r nrow(emu)` documents with a total number of N = `r length(emu_words$word)` words and a vocabulary of V = `r length(unique(emu_words$word))` words. After the exclusion of words with a frequency lower than 5, the vocabulary was reduced to V = `r ncol(emu_dtm)`.

A key tuning parameter in an LDA model is the number of clusters K. A K with a smaller value will yield more general topics whereas a K with a high value will result in more specific topics. Hence, finding a suitable value entails a trade-off between interpretability and specificity. This is very much dependent on the data at hand and the research question. Given the small size of the corpus and the way the program was structured, the value of K = 5 was determined qualitatively considering the input received by the students in the two thematic semesters at TU Delft and the three alternative exchange semesters, each with a very distinct approach:

- TU Delft, Fall semester: Urban region networks
- TU Delft, Spring semester: Constructing the sustainable city
- IUAV Venice, exchange semester:
- KU Leuven, exchange semester:
- UPC Barcelona, exchange semester:

To check how well the clustering is aligned with where the student carried out their exchange semester, we used a confusion matrix to visualize the relation between real values and values assigned through clustering.

@@

@@

```{r lda-k-5}
#| message: false

## K = 5 considering the 2 TUD semesters and the 3 exchange semesters combined 
emu_lda_5 <- LDA(emu_dtm, k = 5, method="Gibbs",
                 control = list(seed = 2023, iter = 500, verbose = 25))

@@

topic_proportions |> 
  left_join(counts_of_primary_topics, by = "topic") |> 
  rename(Topic = topic,
         `Proportion per corpus` = proportion,
         `No. of documents with primary topic` = count)
```
